---
title: "p8105_hw1_mts2203"
author: "Madeline Sharp"
date: "2026-09-19"
output: github_document
---

# Problem 0

## Loading Libraries

``` {r load_libraries, message = F}  

# loading tidyverse library
library(tidyverse)

```


# Problem 1: Coding and Plotting Penguins Dataset

## Loading and Printing Data

``` {r load_penguins, message = F}

# loading penguins dataset from palmerpenguins package
data("penguins", package = "palmerpenguins")

# printing penguins dataset
penguins

``` 

## Short Description

The `palmerpenguin` dataset `penguins` presents a list of `r nrow(penguins)` penguins that are categorized by `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_ (grams)_g`, `sex` and `year`. The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. The penguin's mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

## Creating Scatterplot

``` {r penguins_problem, message = F}

# defining scatterplot of flipper_length_mm vs. bill_length_mm using penguins data
scatterplot_penguins = ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() 

# prining scatterplot
scatterplot_penguins 

# creating and saving scatterplot as a pdf
ggsave("scatterplot_penguins.pdf", height = 4, width = 6)

```


# Problem 2: Variable Types and Coercion

## Defining and Printing Dataframe

```  {r df_and_scatterplot_problem, message = F}

# defining dataframe containing numeric, logical, character and factor vectors
vec_dt_df =
  tibble(
    vec_num = rnorm(n = 10),
    vec_log = vec_num > 0,
    vec_char = c("This", "is","my", "first", "homework", "assignment", "for", "Data", "Science", "I"),
    vec_fact = factor(c("YES", "NO","MAYBE", "YES", "NO", "MAYBE", "YES", "NO", "MAYBE", "YES"))
  )

# printing dataframe containing numeric, logical, character and factor vectors
vec_dt_df

```

## Calculating Dataframe Vector Means

```  {r calculating_vec_means, message = F}

mean_vec_num = mean(pull(vec_dt_df, vec_num))
mean_vec_log = mean(pull(vec_dt_df, vec_log))
mean_vec_char = mean(pull(vec_dt_df, vec_char))
mean_vec_fact = mean(pull(vec_dt_df, vec_fact))

mean_vec_num
mean_vec_log
mean_vec_char
mean_vec_fact

```

## What Works? What Doesn't?

The mean calculations worked for the numeric variable (`vec_num`) and the logical variable (`vec_log`).
The mean calculations did not work for the character variable (`vec_char`) and the factor variable (`vec_fact`).

## Applying `as.numeric`

```  {r applying_as.numeric, results = hide, message = F}

as.numeric(pull(vec_dt_df, vec_log))
as.numeric(pull(vec_dt_df, vec_char))
as.numeric(pull(vec_dt_df, vec_fact))

```

## Explaining `as.numeric` Results

