---
title: "p8105_hw1_mts2203"
author: "Madeline Sharp"
date: "2026-09-19"
output: github_document
---

# Problem 0

## Loading Libraries

``` {r load_libraries, message = F}  

# loading tidyverse library
library(tidyverse)

```


# Problem 1: Coding and Plotting Penguins Dataset

## Loading and Printing Data

``` {r load_penguins, message = F}

# loading penguins dataset from palmerpenguins package
data("penguins", package = "palmerpenguins")

# printing penguins dataset
penguins

``` 

## Short Description

The `palmerpenguin` dataset `penguins` presents a list of `r nrow(penguins)` penguins that are categorized by `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_ (grams)_g`, `sex` and `year`. The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. The penguin's mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

## Creating Scatterplot

``` {r penguins_problem, message = F}

# defining scatterplot of flipper_length_mm vs. bill_length_mm using penguins data
scatterplot_penguins = ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() 

# prining scatterplot
scatterplot_penguins 

# creating and saving scatterplot as a pdf
ggsave("scatterplot_penguins.pdf", height = 4, width = 6)

```


# Problem 2: Variable Types and Coercion

## Defining and Printing Dataframe

```  {r df_and_scatterplot_problem, message = F}

# defining dataframe containing numeric, logical, character and factor vectors
vec_dt_df =
  tibble(
    vec_num = rnorm(n = 10),
    vec_log = vec_num > 0,
    vec_char = c("This", "is","my", "first", "homework", "assignment", "for", "Data", "Science", "I"),
    vec_fact = factor(c("YES", "NO","MAYBE", "YES", "NO", "MAYBE", "YES", "NO", "MAYBE", "YES"))
  )

# printing dataframe containing numeric, logical, character and factor vectors
vec_dt_df

```

## Calculating Dataframe Vector Means

```  {r calculating_vec_means, message = F}

# calculating the mean of vec_num from vec_dt_df
mean(pull(vec_dt_df, vec_num))

# calculating the mean of vec_log from vec_dt_df
mean(pull(vec_dt_df, vec_log))

# calculating the mean of vec_char from vec_dt_df
mean(pull(vec_dt_df, vec_char))

# calculating the mean of vec_fact from vec_dt_df
mean(pull(vec_dt_df, vec_fact))

```

## Explaining `mean` Results

**What Works? What Doesn't?**

The mean calculations worked for the numeric variable (`vec_num`) and the logical variable (`vec_log`).
The mean calculations did not work for the character variable (`vec_char`) and the factor variable (`vec_fact`).

## Applying `as.numeric`

```  {r applying_as.numeric, results = "hide", message = F}

# applying as.numeric to variables of non-numeric vec_log variable from vec_dt_df
as.numeric(pull(vec_dt_df, vec_log))

# applying as.numeric to variables of non-numeric vec_char variable from vec_dt_df
as.numeric(pull(vec_dt_df, vec_char))

# applying as.numeric to variables of non-numeric vec_fact variable from vec_dt_df
as.numeric(pull(vec_dt_df, vec_fact))

```

## Explaining `as.numeric` Results

**What happens? Why?** 

The function `as.numeric` converts values from non-numeric data types to the numeric data type. 

* Logical vector (`vec_log`): all instances of "TRUE" are converted to "1" and all instances of "FALSE" are converted to "0". R reads these "1"s and "0"s as meaningfully numeric since they communicate the proportion of "TRUE" values.

* Character vector (`vec_char`): all 10 elements are converted to "NA" from their assigned character values. R cannot interpret arbitrary text as numbers. However, if a character contains a number as text, R converts this to its corresponding numeric value.

* Factor vector (`vec_fact`): all instances of "YES" are converted to "3", all instances of "NO" are converted to "2" and all instances of "MAYBE" are converted to "1". R converts factor variables into the numeric representations associated with their underlying integer codes. These integer codes are ordered based on the element's level, e.g. "YES" = "3"; "NO" = 2; "MAYBE" = 1. As such, these conversions hold no meaningful numeric weight; they function categorically.

**Does this help explain what happens when you try to take the mean?** 

* Yes, this process communicates that to successfully take the mean of a variable, the variable must either have a numeric data type, or have the capacity to become meaningfully numeric. In this exercise, this includes:
  * the numeric vector, `vec_num`, 
  * and the logical vector, `vec_log`.